---
title: 'Reproducible Research: Peer Assessment 1'
author: "Randy Sherrod"
date: "December 7, 2015"
output: html_document
---
###This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

##Loading and preprocessing the data
```{r read}
setwd("C:\\Training\\RR\\data")
data <- read.csv("activity.csv")
summary(data)
```
##What is mean total number of steps taken per day?
```{r summary}
#Calculate the total number of steps taken per day
steps_per_day <- aggregate(steps ~ date, data, sum)
#Make a histogram of the total number of steps taken each day
hist(steps_per_day$steps,breaks=10, main="Steps per Day", xlab="Steps", col="grey")
#calculate Mean/mediam
mean <- mean(steps_per_day$steps, na.rm=TRUE)
median <- median(steps_per_day$steps, na.rm=TRUE)
```
The mean steps per day is `r mean`.  The median steps per day is `r median`.

##What is the average daily activity pattern?
```{r pattern}
steps_interval <- aggregate(data$steps,by = list(interval = data$interval),
                            data = data, FUN=mean, na.rm=TRUE)
names(steps_interval) <- c("interval","steps")
plot(steps ~ interval, data = steps_interval, type = "l", xlab = "Time Intervals (5-minute)", 
     ylab = "Mean number of steps taken", main = "Average number of steps taken at 5 minute intervals",  col= "black")

##Maximum steps by interval
maxSteps <- max(steps_interval$steps)
##Which interval contains the maximum average number of steps
max_interval <- steps_interval[steps_interval$steps==maxSteps,1]
```
`r max_interval` is the 5-minute interval that contains the maximum number of steps (`r maxSteps`).

##Imputing missing values

```{r impute}
##Number of NAs in original data set
missing <- nrow(data[is.na(data$steps),])

#impute missing values using mean of interval
names(steps_interval) <- c("interval","mean_steps")
w_imputed <- merge(data,steps_interval, by="interval", all.x=TRUE)
w_imputed$steps[is.na(w_imputed$steps)]=w_imputed$mean_steps[is.na(w_imputed$steps)]

steps_interval2 <- aggregate(steps ~ interval, data = w_imputed, sum, na.rm = TRUE)
hist(steps_interval2$steps,breaks=10, main="Steps per Day with Imputed Values", xlab="Steps", col="grey")

#mean with NA's imputed
mean <- mean(steps_interval2$steps, na.rm=TRUE)

#median with NAs imputed
median <- median(steps_interval2$steps, na.rm=TRUE)
```
The number of missing values in the dataset is: `r missing`.

After imputing, the mean steps per day is `r mean`.  The median steps per day is `r median`.

##Are there differences in activity patterns between weekdays and weekends?
```{r days}
w_imputed$day <- ifelse(as.POSIXlt(as.Date(w_imputed$date))$wday%%6 == 0, "weekend", "weekday")
w_imputed$day <- factor(w_imputed$day, levels = c("weekday", "weekend"))
steps.interval= aggregate(w_imputed$steps, by=list(w_imputed$interval,w_imputed$day), data = w_imputed, FUN=mean)
names(steps.interval) <- c("interval","day","steps")
library(lattice)
xyplot(steps ~ interval | factor(day), data = steps.interval, aspect = 1/2, type = "l")
```

The plot clearly shows differences in activity patterns between weekdays and weekends.
